---
title: "Intro to R Lesson 2: Data Transformation"
author: Natanop Pimonsathian
---
Here we are at the Lesson 2. You are about to learn how to transform data, including filtering, indexing, renaming, etc. We will explore a few example of data sets.

The first data set, 'house.txt', contains property sales records between 2006 and 2010. The data has many attributes, such as sale price, sizes, and conditions.   
---
```{r read in data}
houses <- read.delim("~/Dropbox/Gitbucket/Course/input/houses.txt",stringsAsFactors = FALSE)
```

```{r read in data using a relative path}
## reading in 'houses' using a relative path
houses <- read.delim("input/houses.txt", stringsAsFactors = FALSE)
```

Primary Data Exploration 
```{r print the entire dataset}
houses

```
 
```{r displaying Top/Last 5 rows of the data }
head(houses)
tail(houses)
```

Inspect the overall data structure with str()
```{r}
str(houses)
```

Quick question: how many rows are in the 'houses' data set?
```{r displaying the dimensions}
dim(houses)
```
Transforming data using base R 
```{r check if there are missing data points}
is.na(houses)
is.na(houses$PID)
```

Exercise 1: Inspect NAs in columns 'Pool.QC', 'Year.Built', and 'Fence'
```{r Exercise 1 Answers}
#insert your answers here


```
Using the Base R Package, we can do a number of operations on the data. The first example is about creating a new column and deleting some columns
```{r Creating a new column}
houses$Total.Lot.Area <- houses$Lot.Frontage + houses$Lot.Area
```
```{r Deleting irrelevant columns}
houses$Order <- NULL 
```
Exercise 2: Creating a new column called 'Overall.rating' as an average score between 'Overall.Qual' and 'Overall.Cond'. 
```{r Exercise 2 Answers}
# insert your answer here

```
We can use the Base R package to filter the numbers of observations
```{r Filter to see the first 4 houses}
four_houses <- houses[c(1, 2, 3, 4), ]
dim(four_houses)
```
```{r Filter to see the first 10 houses}
ten_houses <- houses[1:10, ]
dim(ten_houses)
```

Selecting particular rows based on their properties. The below operations are used to filter the rows that fall under the conditions

* `>`: greater than
* `>=`: greater than or equal to
* `<`: less than
* `<=`: less than or equal to
* `==`: equal to
* `!=`: not equal to
* `!`: not
* `&`: and
* `|`: or

Example: Filter the houses that are cheaper than 100k 
```{r subsetting cheap houses}
cheap_houses <- houses[houses$SalePrice < 100000, ]
```
Find the houses that were built and remodified in the same year (Year.Built = Year.Remod.AA)
```{r non-modified houses}
n_mod_houses <- houses[houses$Year.Built == houses$Year.Remod.Add, ]
```
select all houses built before 1900, storing only the basement columns.
```{r old basement columns}
old_basements <- houses[houses$Year.Built < 1900,
  grepl("Bsmt", names(houses))]
```
The 'grepl' command is used to check whether the first argument is a substring of the second argument 
```{r Garage columns for houses with a big garage}
garage_big_houses <- houses[houses$Garage.Area >= 400 ,
  grepl("Garage", names(houses))]

```
Also, we can use '%in%' to check whether the left input is in the right input 
```{r %in% demonstration}
c("Risa", "Ham", "Bow", "Jane") %in% c("Ham", "Jane")
```


