---
title: "Intro to R Lesson 3: Data Transformation Part 2"
author: Natanop Pimonsathian
---
This lesson is the continued story of Lesson 2. Last week, we learned to import the data file into R environment and to use some functions on data manipulation. This week, we will be learning to process the data with a more intelligent way. 

Let's recall a few things from the previous unit. We have used to filter data with operations and functions in Base R. Those tools have a number of advantages and also a bit of disadvantages. 

Remember the last exercise from the previous chapter? Where we had to extract the observations that meet the criteria. We have to write several lines of code to archive that. 

```{r read in data using a relative path}
## reading in 'houses' using a relative path
houses <- read.delim("input/houses.txt", stringsAsFactors = FALSE)
```

Refreshing Exercise: Name all the houses that
- Built after the year 1990 
- Have overall quality rating more than or equal to 7.00
- Are accessible by paved street 
- Selling prices are less than 200000 USD 
- Extract the columns that are related to basement only 

```{r recall the exercise}
houses_subset <- houses[houses$Year.Built > 1990 &
                        houses$Overall.Qual >= 7 &
                        houses$Street == 'Pave'& 
                        houses$SalePrice < 200000,
                        grepl("Bsmt", names(houses))]
```

```{r}
houses_subset_2 <- houses %>%
    filter(Year.Built > 1990, Overall.Qual >= 7, 
           Street == 'Pave', SalePrice < 200000)
  
```

```{r select columns that contain Bsmt with dplyr}
houses_Bsmt <- houses %>% 
  select(contains("Bsmt"))
```



Now we will learn to use the Pipe operator, %>%  
To use the operator, we need to install the package called 'dplyr'. 

```{r install dplyr}
install.packages('dplyr')
library('dplyr')

```

Let's perform a test on the basic operation: filtering the houses that build after the year 1990 and have overall quality more than or equal to 7. 

Performing this with BaseR will look like
```{r Base R demo}
houses_old_good1 <- houses[houses$Year.Built > 1990 &
                          houses$Overall.Qual >= 7, ]
dim(houses_old_good1)
```

Performing this with dplyr will look like
```{r dplyr demo}
houses_old_good2 <- houses %>%
  filter(Year.Built > 1990, Overall.Qual >= 7)

dim(houses_old_good2)
```

What we can do with the Pipe operation?
1. Filter 
```{r Filter demo: find the houses with a large garage}
houses_large_grg <- houses %>%
  filter(Garage.Area > 300)

one_family_dplyr <- houses %>%
  filter(Bldg.Type == "1Fam")
```

2. Select some columns
```{r Select columns demo: extracting sale prices and year built}
houses_price_year <- houses %>%
  select(SalePrice, Year.Built)

dim(houses_price_year)
```
```{r}
houses_cleaned <- houses %>%
  select(-Order, -SalePrice)

```

3. Create a new column 
```{r Create a new column}
houses_overall_rating <- houses %>% 
  mutate(Overall_rating = (Overall.Cond+Overall.Qual)/2)
```

4. Sort rows  
```{r Sorting }
## sorting selling prices from high to low 
houses_price <- houses %>%
  select(PID, SalePrice)

sorted_prices_low <- arrange(houses_price, SalePrice) #from low to high
sorted_prices_high <- arrange(houses_price, desc(SalePrice)) #from high to low
```

Exercise: Select columns: select the column that contains "Garage" in their names. Then sort the area from the largest to the smallest

```{r Select columns and sort }
garage_area <- houses %>%
  select(Garage.Area) %>%
  arrange(desc(Garage.Area))

```

5. Grouping
```{r Grouping}
## adding a new grouping based on two variables &
## summarizing the mean sales prices
price_per_type <- houses %>%
  group_by(Bldg.Type) %>%
  summarise(Mean.Price = mean(SalePrice))
```

Exercise: Summarize the number of houses and median price per building type. 
```{r No. of houses and thier median prices per building type}
price_per_type <- houses %>%
  group_by(Bldg.Type) %>%
  summarise(Med.Price = median(SalePrice))
```



