---
Title: "Intro to R Lesson 4: Data Visualization"
Author: Natanop Pimonsathian
Instructor: Natanop Pimonsathian
---
This lesson discusses the common visualization tools. We will learn to transform
a complex Data Frame into an intelligent pictorial presentation. 

We will focus on
- The elements of a visualization
- Plotting Scatterplots, boxplots, barplots, lineplots
- Working with the Grammar of Graphics (ggplot2) library. 

```{r read in data using a relative path}
## reading in 'houses' using a relative path
houses <- read.delim("input/houses.txt", stringsAsFactors = FALSE)
```

Plotting in one dimension - Boxplot

```{r boxplot demo}
boxplot(houses$SalePrice)
```
Exercise 1: plot a boxplot of the variable 'Overall.Qual'
```{r Exercise 1 Answer}
boxplot(houses$Overall.Qual)
```

Plotting in two dimensions - Scatterplot
```{r boxplot Scatterplot}
plot(x=houses$Overall.Qual,y=houses$SalePrice)
plot(y=houses$Overall.Qual,x=houses$SalePrice)
```
Exercise 2.1: plot a scatterplot of variables 'SalePrice' as a function of 'Overall.Cond'
```{r Exercise 2.1 Answer}
plot(x=houses$Overall.Cond,y=houses$SalePrice)

```

Exercise 2.2: plot a scatterplot of variables 'Overall.Qual' as a function of 'Overall.Cond'
```{r Exercise 2.2 Answer}
plot(x=houses$Overall.Cond,y=houses$Overall.Qual)

```

Now, you may wonder how can we adjust the aesthetic of the plot. For example,
how to add the title of the plot. The below chunk shows a full modification of 
a scatter plot. 
```{r Full plot demo}
plot(x=houses$Overall.Qual, y=houses$SalePrice, 
     main = "Price as function of the quality score",
     xlab = "Quality Score", ylab = "Selling Price",
     pch = 19)
```

What if you want to decorate more? We can use the so-called ggplot2 library. 
```{r installing the ggplot2}
install.packages("ggplot2")
library(ggplot2)
```

The first demonstration is to plot the same plot as shown before, the association
between house quality and selling price. 
```{r ggplot2 First Demo}
ggplot() +
  geom_point(data = houses, mapping = aes(x = Overall.Qual, y = SalePrice))
```

We will add another element, a regression line, in order to the see the trend.  
```{r Scatterplot with the best fit line}
ggplot() +
  geom_point(data = houses, aes(x = Overall.Qual, y = SalePrice)) +
  geom_smooth(data = houses, aes(x = Overall.Qual, y = SalePrice))
```

This can also written in a more compact format.  
```{r Scatterplot with the best fit line - alternate version}
ggplot(data = houses, aes(x = Overall.Qual, y = SalePrice)) +
  geom_point() +
  geom_smooth()
```

Exercise 3: plot a scatterplot of variables 'Overall.Qual' as a function of 'Overall.Cond' and also plot the best fit line. 

```{r Exercise 3 Answer }
ggplot(data=houses, mapping = aes(x=Overall.Cond, y=Overall.Qual)) +
  geom_point()
  #geom_smooth()

```

Plotting a boxplot in ggplot2 
```{r Boxplot demo}
ggplot(data = houses, aes(y = SalePrice)) +
  geom_boxplot()
```

How can we compare between different building conditions? We may add the 
x-axis to group the selling price into each building type. 
```{r Boxplot demo 2}
ggplot(data = houses, aes(x = Bldg.Type, y = SalePrice)) +
  geom_boxplot()
```
We can also add color into each chunk 
```{r Scatterplot with color demo}
ggplot(data = houses, aes(x = Overall.Qual, y = SalePrice, color = Bldg.Type,
                          shape = Bldg.Type, size = Bldg.Type)) +
  geom_point()
```
Barplot
```{r Barplot Demo}
ggplot(data = houses, aes(x = Bldg.Type)) +
  geom_bar()
```
Histogram
```{r Histogram demo}
ggplot(data = houses, aes(x = Gr.Liv.Area)) +
  geom_histogram(bins = 100)
```
Line plots: plotting with statistics

```{r Stat count}
ggplot(data = houses, aes(x = Yr.Sold)) +
  geom_line(stat = "count")
```
```{r Stat mean}
ggplot(data = houses, aes(x = Bldg.Type, y = SalePrice)) +
  geom_bar(stat = "summary")
```
Additional things to know: Chart Decoration 
```{r Naming things}
plot1 <- ggplot(data = houses, aes(x = Bldg.Type, y = SalePrice)) +
  geom_bar(stat = "summary")

plot1.1 <- plot1 + ggtitle("Revenue per bulding type") +
  xlab("Property type") + 
  ylab("Total Revenue")
```

```{r}
plot1 + ggtitle("Revenue per bulding type") +
  xlab("Property type") + 
  ylab("Total Revenue")

plot1
```

```{r}
plot2 <- ggplot(data = houses, aes(x = Lot.Area, y = SalePrice, color = House.Style,
                          size = Overall.Qual)) +
  geom_point() +
  ggtitle("Revenue per lot area") +
  xlab("Lot Area") +
  ylab("Selling Price")
```

